### 2019-08-26 manuscript.tex
update

\documentclass[11pt,a4paper]{article}

%%%%%%%%%%%%%%%%%%%%%%%
%%% IMPORT PACKAGES %%%
%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[margin=2.5cm,a4paper]{geometry}

% text packages
\usepackage{csquotes}
\usepackage[english]{babel}
\usepackage{setspace}

% scripting packages
\usepackage{blindtext}

% style packages
\usepackage{lmodern}
\usepackage{fancyhdr}
\usepackage{microtype} % Sublim­i­nal re­fine­ments to­wards ty­po­graph­i­cal per­fec­tion

% figure packages
\usepackage{tikz}
\usetikzlibrary{calc,shapes,arrows,intersections,shadows}
\tikzset{>=latex}
\usepackage{graphicx}
\usepackage[indention=0.5cm,labelsep=colon,font={sf,small},labelfont={bf,sf}]{caption}
\usepackage[indention=0.5cm,font={sf,small},labelfont={bf,sf}]{subcaption}
\usepackage{float}
\usepackage{multirow}
\usepackage{import}
\usepackage{xcolor}
\usepackage{colortbl}

% pseudocode packages
\usepackage{algorithm}
\usepackage{algpseudocode}

% symbols packages
\usepackage{amssymb}
\usepackage{amsmath}

% table packages
\usepackage{tabularx}
\usepackage{tabulary}
\usepackage{booktabs}

% references packages
\usepackage[sorting=none]{biblatex}
\bibliography{library}
\usepackage{bibentry}

% link packages
\usepackage{url}
\usepackage{doi}

% add checkmark commands
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% DEFINE FIGURE HELPERS %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\hugefigure}{\textwidth}
\newcommand{\LARGEfigure}{0.8\textwidth} % 1 per col
\newcommand{\LARgefigure}{0.6\textwidth} % 1 per col
\newcommand{\Largefigure}{0.5\textwidth} % 1 per col
\newcommand{\largefigure}{0.37\textwidth} % 2 per col
\newcommand{\normalfigure}{0.3\textwidth} % 3 per col
\newcommand{\smallfigure}{0.24\textwidth} % 3 per col
\newcommand{\Smallfigure}{0.19\textwidth} % 4 per col
\newcommand{\SMALLfigure}{0.15\textwidth} % 5 per col

%%%%%%%%%%%%%%%%%%%%%
%%% DEFINE STYLES %%%
%%%%%%%%%%%%%%%%%%%%%

\usepackage{fontspec}
\setmainfont [Path = fonts/,
UprightFont = *-300,
ItalicFont = *-300-Italic,
BoldFont = *-700,
BoldItalicFont = *-700-Italic
]{MuseoSans}

%% hyperlinks
\usepackage{hyperref}

\renewcommand{\baselinestretch}{1.25}

\begin{document}

\title{dyngen: simulating developing single cells}

\author{The \textit{dynverse} guys}

\maketitle

\input{content}

\appendix

\printbibliography

\end{document}