### 2019-11-18 analysis/vignette_network_inference.Rmd
commit derivative code for GRN NI

---
title: "Use-case: evaluation of network inference"
author: "Robrecht Cannoodt"
output: rmarkdown::html_vignette
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{Use-case: evaluation of network inference}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dyngen)

set.seed(1)

backbone <- backbone_bifurcating_cycle()
num_kos <- 40
model <- 
  initialise_model(
    num_cells = 1000,
    num_tfs = 200,
    num_targets = 900,
    num_hks = 900,
    distance_metric = "pearson",
    backbone = backbone,
    tf_network_params = tf_network_default(min_tfs_per_module = 2, sample_num_regulators = function() 2),
    feature_network_params = feature_network_default(target_resampling = 5000),
    kinetics_params = kinetics_default(),
    gold_standard_params = gold_standard_default(),
    simulation_params = simulation_default(
      census_interval = 1,
      burn_time = simtime_from_backbone(backbone, TRUE),
      total_time = simtime_from_backbone(backbone),
      experiment_params = bind_rows(
        simulation_type_wild_type(num_simulations = 32),
        simulation_type_knockdown(num_simulations = num_kos, num_genes = sample(1:10, num_kos, replace = TRUE))
      ),
      store_reaction_propensities = TRUE,
      store_grn = TRUE
    ),
    experiment_params = experiment_snapshot(),
    verbose = TRUE,
    download_cache_dir = "~/.cache/dyngen",
    num_cores = 8
  )

model <- model %>% 
  generate_tf_network() %>% 
  generate_feature_network() %>% 
  generate_kinetics() %>% 
  generate_gold_standard() %>%
  generate_cells() %>% 
  generate_experiment()

traj <- 
  model %>% 
  wrap_dataset(store_grn = TRUE)

traj$expression

reactions <- model$simulation_system$reactions 
names(reactions) <- map_chr(reactions, "name")

comp_reac <- dyngen:::.generate_cells_precompile_reactions(model)
sim <- dyngen:::.generate_cells_simulate_cell(
  simulation_i = 1, model, comp_reac, debug = TRUE
)

feature_network <- model$feature_network

knockouts <- 
  bind_rows(pbapply::pblapply(
    seq_len(nrow(feature_network)), 
    function(i) {
      from <- feature_network$from[[i]]
      to <- feature_network$to[[i]]
      reac_ix <- match(paste0("transcription_", to), comp_reac$reaction_ids)
      from_ix <- match(paste0("y_", from), names(sim$initial_state))

      sim$state <- model$simulations$counts[traj$cell_info$step_ix[[1]],] %>% unname
      sim$calculate_propensity()

      orig_prop <- sim$propensity[[reac_ix]]

      sim$state[[from_ix]] <- 0
      sim$calculate_propensity()
      new_prop <- sim$propensity[[reac_ix]]

      tibble(
        i,
        from,
        to,
        orig_prop,
        new_prop,
        reg_effect = orig_prop - new_prop,
        buffer_val = traj$feature_network_sc[1,paste0("regulation_", from, "_", to)]
      )
    }
  ))
max_prot <- model$feature_info %>% select(feature_id, max_protein) %>% deframe()
wpr <- model$feature_info %>% select(feature_id, wpr) %>% deframe()
xdr <- model$feature_info %>% select(feature_id, xdr) %>% deframe()
ggplot(knockouts) +
  geom_point(aes(log10(buffer_val), reg_effect / (wpr[to] - xdr[to]), colour = feature_network$effect %>% as.character)) +
  theme_bw()

feature_network$derivative <- 
  pbapply::pblapply(seq_len(nrow(feature_network)), function(i) {
  from <- feature_network$from[[i]]
  to <- feature_network$to[[i]]

  reaction <- reactions[[paste0("transcription_", to)]]
  prop <- gsub("pow\\(([^,]*) *, *([^\\)]*)\\)", "(\\1)^(\\2)", reaction$propensity)
  props <- prop %>% strsplit(" *; *") %>% first()
  propone <- props %>% last()
  subs <- tibble(
    var = gsub(" *=.*", "", props[-length(props)]),
    val = gsub(".*= *", "", props[-length(props)])
  )
  for (j in seq_len(nrow(subs))) {
    propone <- gsub(subs$var[[j]], subs$val[[j]], propone, fixed = TRUE)
  }

  deriv <- D(parse(text = propone), paste0("y_", from))
})

state <- c(
  model$simulations$counts[traj$cell_info$step_ix[[1]],],
  model$simulation_system$parameters
)
deriv <- feature_network$derivative[[1]]
# list2env(as.list(state), .GlobalEnv)
# eval(deriv)
env <- new.env()
list2env(as.list(state), env)

derivs <- map_dbl(feature_network$derivative, ~ eval(., envir = env))
reac_props <- model$simulations$reaction_propensities[traj$cell_info$step_ix[[1]], paste0("transcription_", model$feature_network$to)]

qplot(
  log10(traj$feature_network_sc[1,]),
  derivs,
  colour = reac_props
  # colour = as.character(feature_network$effect)
) + labs(
  x = "log10 sc importance = log10((y/k)^c)",
  y = "deriv. propensity function", 
  colour = "effect"
) +
  theme_bw() +
  viridis::scale_colour_viridis()
ggsave("~/deriv.pdf", width = 10, height = 10)

qplot(
  log10(traj$feature_network_sc[1,]),
  knockouts$reg_effect,
  colour = as.character(feature_network$effect)
) + labs(
  x = "log10 sc importance = log10((y/k)^c)",
  y = "KO", 
  colour = "effect"
) +
  theme_bw()

qplot(
  reac_props,
  derivs
) +
  theme_bw()

qplot(
  reac_props,
  log10(traj$feature_network_sc[1,])
) +
  theme_bw()
```